---
title: "STA2102 Homework 1"
author: "Kevin Zhang 1002225264"
date: \today
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

# a)

\[\frac{H_m\hat{Z}H_n}{mn}=\frac{H_mH_mZH_nH_n}{mn}=\frac{mIZnI}{mn}=Z\]
Notice that $H_m$ and $H_n$ are symmetric and $H_mH_m^\intercal=mI$ and $H_nH_n^\intercal=nI$.

# b)

```{r}
hard_th <- function(x, lambda){
  I = nrow(x)
  J = ncol(x)
  for(i in 1:I){
    for(j in 1:J){
      x[i,j] = ifelse(abs(x[i,j])>lambda, x[i,j], 0)
    }
  }
  return(x)
}
soft_th <- function(x, lambda){
  I = nrow(x)
  J = ncol(x)
  for(i in 1:I){
    for(j in 1:J){
      x[i,j] = sign(x[i,j]) * max(abs(x[i,j])-lambda, 0)
    }
  }
  return(x)
}
fwht2d <- function(x){
  h <- 1
  len <- ncol(x)
  while(h < len){
    for(i in seq(1,len,by=h*2)){
      for (j in seq(i,i+h-1)){
        a <- x[,j]
        b <- x[,j+h]
        x[,j] <- a + b
        x[,j+h] <- a - b
      }
    }
    h <- 2*h
  }
  h <- 1
  len <- nrow(x)
  while(h < len){
    for(i in seq(1,len,by=h*2)){
      for(j in seq(i,i+h-1)){
        a <- x[j,]
        b <- x[j+h,]
        x[j,] <- a + b
        x[j+h,] <- a - b
      }
    }
    h <- 2*h
  }
  x
}
```

# c)

```{r}
design <- matrix(scan("design.txt"),ncol=256,byrow=T)
colours <- grey(seq(0,1,length=256))
image(design, axes=F, col=colours)
xhat <- fwht2d(design)
for(i in c(50, 100, 200)){
  xhat_soft <- soft_th(xhat, i)
  xhat_hard <- hard_th(xhat, i)
  x_soft <- fwht2d(xhat_soft)/ncol(xhat_soft)^2
  x_hard <- fwht2d(xhat_hard)/ncol(xhat_hard)^2
  image(x_soft, axes=F, col=colours, main=sprintf('soft thresholding: lambda = %d', i))
  image(x_hard, axes=F, col=colours, main=sprintf('hard thresholding: lambda = %d', i))
}
```

## Question 2

# a)

\[E(t^S)=\sum_{n=0}^\infty E(t^X|N=n)P(N=n)=\sum_{n=0}^N(\phi(t))^n(1-\theta)\theta^n=\frac{1-\theta}{1-\theta\phi(t)}=g(\phi(t))\]

Note that $\sum_{k=0}^\infty ar^k=\frac{a}{1-r}$.

# b)

Suppose $N<m$, then $S=\sum_{i=1}^N X_i\leq\sum_{i=1}^{m-1} X_i\leq(m-1)l<ml$. Therefore, by contrapositive, $S\geq ml$ implies $N\geq m$. Hence, $P(S\geq ml)\leq P(N\geq m)\leq\epsilon$.

# c)

\begin{gather*}
  P(S\geq M)\leq\frac{1}{t^M}\left(\frac{1-\theta}{1-\theta\phi(t)}\right)\implies P(S\geq M)\leq\inf_{1<\phi(t)<\theta^{-1}}\frac{1}{t^M}\left(\frac{1-\theta}{1-\theta\phi(t)}\right)\\
  \text{suppose }\inf_{1<\phi(t)<\theta^{-1}}\frac{1}{t^M}\left(\frac{1-\theta}{1-\theta\phi(t)}\right)=\epsilon\text{ i.e. }P(S\geq M)\leq\epsilon\\
  \text{since }\ln(x)\text{ is continuous, then }\inf_{1<\phi(t)<\theta^{-1}}\ln(1-\theta)-\ln(1-\theta\phi(t))-M\ln(t)=\ln(\epsilon)\\
  \implies \inf_{1<\phi(t)<\theta^{-1}}\frac{\ln(1-\theta)-\ln(1-\theta\phi(t))-\ln(\epsilon)}{\ln(t)}=M
\end{gather*}

# d)

```{r}
p <- numeric(0)
for(i in 0:10){
  p <- c(p, choose(10, i)*0.5^10)
}
phi <- function(t){
  return((1+t)^10/2^10)
}
get_dist_s <- function(x_max, epsilon, phi, theta){ # x_max = l
  t_max <- (1/0.9*2^10)^0.1-1 # the max value for t such that phi(t)<1/theta
  t_seq <- seq(1, t_max, by=(t_max-1)/100)
  t_seq <- t_seq[-1] # remove t=1
  t_seq <- t_seq[-length(t_seq)] # remove t_max
  M <- Inf
  for(t in t_seq){
    M_temp <- (log(1-theta)-log(1-theta*phi(t))-log(epsilon))/log(t)
    if(M_temp < M){
      M <- ceiling(M_temp)
    } # M = min of the function
  }
  p_dft <- numeric(0) # dft of all the p
  for(j in 0:(M-1)){
    p_temp <- 0
    for(x in 0:x_max){
        p_temp <- p_temp + exp(-2*pi*complex(real=0, imaginary=1)*j/M*x)*p[x+1]
    }
    p_dft <- c(p_dft, p_temp)
  }
  g <- numeric(0) # g of p
  for(p_j in p_dft){
    g <- c(g, (1-theta)/(1-theta*p_j))
  }
  ps <- numeric(0) # vector of p(S=s)
  for(s in 0:(M-1)){
    ps_temp <- 0
    for(j in 0:(M-1)){
      ps_temp <- ps_temp + exp(2*pi*complex(real=0,imaginary=1)*s/M*j)*g[j+1]
    }
    ps <- c(ps, round(Re(ps_temp/M), 8))
  }
  return(list(ps, M)) #jth element of ps is p(S=j-1)
}
```

```{r}
t_max <- (1/0.9*2^10)^0.1-1 # the max value for t such that phi(t)<1/theta
t_seq <- seq(1, t_max, by=(t_max-1)/100)
t_seq <- t_seq[-1] # remove t=1
t_seq <- t_seq[-length(t_seq)] # remove t_max
x_max <- 10
epsilon <- 10^-5
theta <- 0.9
plot(t_seq, (log(1-theta)-log(1-theta*phi(t_seq))-log(epsilon))/log(t_seq),
     ylab='function to minimize', type='h')
M <- get_dist_s(x_max, epsilon, phi, theta)[[2]]
ps <- get_dist_s(x_max, epsilon, phi, theta)[[1]]
abline(h=M, col='red')
text(x=1.02, y=5*M, label=sprintf('M is %d', M))
plot(0:(M-1), ps, type='h', xlab='S', ylab='prob')
```


















































































