---
title: "STA2102 Homework 3"
author: "Kevin Zhang 1002225264"
date: "11/21/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

## a)

$$E((V^\intercal AV)^2)=\sum_{i=1}a_{ii}^2 E(V_i^4)=\sum_{i=1}^n\sum_{j=1}^n a_{ii}a_{jj}E(V_i^2)E(V_j^2)=\sum_{i=1}^n\sum_{j=1}^n a_{ii}a_{jj}$$

Therefore we need to minimize $E(V_i^4)$. Note that $E(V_i^4)=var(V_i^2)+(E(V_i^2))^2=var(V_i^2)+1$ is minimized if $var(V_i^2)=0$ or equivalently $V_i^2$ is degenerate at some number $k$. Since $E(V_i^2)=1$, we have $V_i^2=1$ a.s. which implies that $V_i\in\{-1, 1\}$ and $P(-1)=P(1)=0.5$.

## b)

By simple matrix-vector multiplication:
$$\begin{pmatrix}H_{11}&H_{12}\\H_{21}&H_{22}\end{pmatrix}\begin{pmatrix}V\\0\end{pmatrix}=\begin{pmatrix}H_{11}V\\H_{21}V\end{pmatrix}$$

$$\begin{pmatrix}H_{11}&H_{12}\\H_{21}&H_{22}\end{pmatrix}\begin{pmatrix}H_{11}^{k-1}V\\0\end{pmatrix}=\begin{pmatrix}H_{11}H_{11}^{k-1}V\\H_{21}H_{11}^{k-1}V\end{pmatrix}=\begin{pmatrix}H_{11}^k V\\H_{21}H_{11}^{k-1}V\end{pmatrix}$$

## c)

```{r}
leverage <- function(x,w,r=10,m=100){
  set.seed(12345) # modification
  qrx <- qr(x)
  n <- nrow(x)
  lev <- NULL
  for(i in 1:m){
    v <- ifelse(runif(n)>0.5,1,-1)
    v[-w] <- 0
    v0 <- qr.fitted(qrx,v)
    f <- v0
    for(j in 2:r){
      v0[-w] <- 0
      v0 <- qr.fitted(qrx,v0)
      f <- f + v0/j
    }
    lev <- c(lev,sum(v*f))
  }
  std.err <- exp(-mean(lev))*sd(lev)/sqrt(m)
  lev <- 1 - exp(-mean(lev))
  r <- list(lev=lev,std.err=std.err)
  r
}
x <- c(1:1000)/1000
X1 <- 1
for (k in 1:5) X1 <- cbind(X1,cos(2*k*pi*x),sin(2*k*pi*x))
library(splines) # loads the library of functions to compute B-splines
X2 <- cbind(1,bs(x,df=10))
plot(x,X2[,2])
for (i in 3:11) points(x,X2[,i])
lev1 <- NULL
err1 <- NULL
lev2 <- NULL
err2 <- NULL
for(k in 1:20){
  w <- (1:1000)[x>(k-1)/20 & x<=k/20]
  r1 <- leverage(X1, w)
  r2 <- leverage(X2, w)
  lev1 <- c(lev1, r1$lev)
  err1 <- c(err1, r1$std.err)
  lev2 <- c(lev2, r2$lev)
  err2 <- c(err2, r2$std.err)
}
data.frame(k=1:20, lev1=lev1, err1=err1, lev2=lev2, err2=err2)
```

The estimates for the two models are not exactly close to each other. The standard error for model 2 seems to be smaller.

# Question 2

## a)

Let $x_m$ denote the median then
\begin{gather*}
  \int_{-\infty}^{x_m} \frac{1}{\pi\sigma}\left(\frac{\sigma^2}{(x-\theta)^2+\sigma^2)}\right)\,dx=\int_{-\infty}^{x_m-\theta} \frac{\sigma}{\pi}\left(\frac{1}{s^2+\sigma^2)}\right)\,ds=\left.\frac{1}{\pi}arctan\frac{s}{\sigma}\right|_{-\infty}^{x_m-\theta}=0.5\\
  \implies\frac{1}{\pi}arctan\frac{x_m-\theta}{\sigma}+0.5=0.5\implies x_m=\theta
\end{gather*}
Similarly, let $x_q$ denote the third quartile, then
$$\left.\frac{1}{\pi}arctan\frac{s}{\sigma}\right|_0^{x_r-\theta}=\frac{1}{4}\implies arctan\frac{x_r-\theta}{\sigma}=\frac{\pi}{4}\implies x_r=\theta+\sigma$$
By symmetry around $\theta$, the IQR is $2\sigma$.

## b)

$$l(x,\theta,\sigma)=n\log\sigma-n\log\pi-\sum_{i=1}^n\log[(x_i-\theta)^2+\sigma^2]$$
$$\frac{\partial l}{\partial\theta}=\sum_{i=1}^n\frac{2(x_i-\theta)}{(x_i-\theta)^2+\sigma^2}$$
$$\frac{\partial l}{\partial\sigma}=\frac{n}{\sigma}-\sum_{i=1}^n\frac{2\sigma}{(x_i-\theta)^2+\sigma^2)}$$
$$\frac{\partial^2l}{\partial\theta^2}=\sum_{i=1}^n\frac{2(x_i-\theta)^2-2\sigma^2}{((x_i-\theta)^2+\sigma^2)^2}$$
$$\frac{\partial^2 l}{\partial\sigma^2}=\frac{-n}{\sigma^2}-\sum_{i=1}^n\frac{2((x_i-\theta)^2+\sigma^2)-4\sigma^2}{((x_i-\theta)^2+\sigma^2)^2}$$
$$\frac{\partial^2l}{\partial\theta\partial\sigma}=\sum_{i=1}^n\frac{4\sigma(x_i-\theta)}{-((x_i-\theta)^2+\sigma^2)^2}$$

```{r}
newton_raph <- function(x, n_iter=1000){
  theta <- median(x)
  sigma <- IQR(x) / 2
  res <- c(theta, sigma)
  h1 <- function(x, theta, sigma){
    temp1 <- 0
    for(x_temp in x){
      temp1 <- temp1 + 2*(x_temp-theta)/((x_temp-theta)^2+sigma^2)
    }
    temp2 <- length(x)/sigma
    for(x_temp in x){
      temp2 <- temp2 - 2*sigma/((x_temp-theta)^2+sigma^2)
    }
    return(c(temp1, temp2))
  }
  h2 <- function(x, theta, sigma){
    temp11 <- 0
    for(x_temp in x){
      temp11 <- temp11 + (2*(x_temp-theta)^2-2*sigma^2)/((x_temp-theta)^2+sigma^2)^2
    }
    temp22 <- -length(x)/sigma^2
    for(x_temp in x){
      temp22 <- temp22 - (2*(x_temp-theta)^2-2*sigma^2)/((x_temp-theta)^2+sigma^2)^2
    }
    temp12 <- 0
    for(x_temp in x){
      temp12 <- temp12 - 4*sigma*(x_temp-theta)/((x_temp-theta)^2+sigma^2)^2
    }
    return(matrix(c(temp11, temp12, temp12, temp22), nrow=2))
  }
  for(i in 1:n_iter){
    h1_temp <- h1(x, res[1], res[2])
    h2_temp <- h2(x, res[1], res[2])
    res <- res - solve(h2_temp) %*% h1_temp
  }
  return(list(res=res, err=-solve((h2(x, res[1], res[2])))))
}
x <- rcauchy(n=1000, location=0,scale=1)
newton_raph(x)
```





















