---
title: "STA2102 Midterm"
author: "Kevin Zhang 1002225264"
date: "10/28/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

```{r}
x <- scan("EV.txt", what=list(0,0))
a <- x[[1]]
theta <- x[[2]]
l <- max(a)
n <- length(a)
t <- numeric(0)
p <- numeric(n*l + 1)
phi_value <- numeric(n*l + 1)
# compute phi(t0) to phi(t_nl)
for(j in 0:(n*l)){
  phi_value_temp <- 1
  for(i in 1:n){
    temp <- 1-theta[i]+theta[i]*exp(-2*pi*complex(imaginary=1)*j*a[i]/(n*l+1))
    phi_value_temp <- phi_value_temp * temp
  }
  phi_value[j+1] <- phi_value_temp
}
# compute p(s=o) to p(s=nl)
for(s in 0:(n*l)){
  p_temp <- numeric(n*l + 1)
  for(j in 0:(n*l)){
    p_temp[j+1] <- phi_value[j+1]*exp(2*pi*complex(imaginary=1)*j*s/(n*l+1))
  }
  p[s+1] <- round(Re(sum(p_temp)/(n*l+1)), 5)
}
p1 <- sum(p[(270+1):(300+1)])
p2 <- p[269+1]
print(sprintf('The answer to part a is %f', p1))
print(sprintf('The answer to part b is %f', p2))
```

# Question 2

## a)

Case 1: $y<0$

\begin{gather*}
  P(U_1-U_2\leq y)=\int_{-y}^1\int_0^{U_2+y}dU_1dU_2=\frac{1}{2}(1-y^2)+y(1-y)\\
  \implies\frac{\partial P(Y\leq y)}{y}=-y+1+2y=1+y=1-|y|
\end{gather*}

Case 2: $y\geq0$

\begin{gather*}
  P(U_1-U_2\leq y)=\int_0^{1-y}\int_0^{U_2+y}dU_1dU_2+\int_{1-y}^1dU_2=\frac{1}{2}(1-y)^2+y(1-y)+y\\
  \implies\frac{\partial P(Y\leq y)}{y}=y-1+1-2y+1=1-y=1-|y|
\end{gather*}

Therefore, $g(y)=1-|y|$ for $y\in[-1,1]$.

## b)

The problem is to find $M$ equal to the max of $\frac{f(x)}{g(x)}$ which is symmetric around 0, so we only look at the interval $[0,1]$. At $x=0$, this evaluate to $\frac{3}{4}$. At $x=1$, by L'Hospital's rule, this evaluates to $\lim_{x\to1}\frac{\frac{3x}{2}}{x}=\frac{3}{2}$. At $x\in(0,1)$, we have
\[\frac{\partial}{\partial x}\left(\frac{f(x)}{g(x)}\right)=\frac{\frac{-3}{2}x(1-x)+\frac{3}{4}(1-x^2)}{(1-x)^2}=0\]
This implies that $\frac{-3}{2}x(1-x)+\frac{3}{4}(1-x^2)=0\implies x^2-2x+1=0\implies x\in\{-1,1\}$. This contradicts with $x\in(0,1)$. Therefore, the max value is equal to $\frac{3}{2}$ at $x\in\{-1,1\}$. Therefore, the acception probability is $\frac{1}{M}=\frac{2}{3}$.

# Question 3

## a)

First, by iterative method, $x_k=\sum_{i=0}^{k-1}B^i b+B^kx_0$. Notice that $B^2=uv^\intercal uv^\intercal=qB$ where $q=v^\intercal u$. Therefore, $B^kx_0=q^{k-1}x_0\to0$ as $k\to\infty$ since $|q|<1$. Now
\begin{gather*}
  (I-B)\sum_{i=0}^{k-1}B^i=\sum_{i=0}^{k-1}B^i-\sum_{i=1}^kB^i=I-B^k\\
  \implies\sum_{i=0}^{k-1}B^i=(I-B)^{-1}(I-B^k)\\
  \sum_{i=0}^{k-1}B^i\to(I-B)^{-1}\text{ since }B^k\to0\text{ or }I-B^k\to I
\end{gather*}
Therefore, $x_k=\sum_{i=0}^{k-1}B^i b+B^kx_0\to(I-B)^{-1}b=x^*$.

## b)

```{r}
B <- 0.1 * matrix(c(-1, -1,  1, -2, -3,
                    -1,  1, -1,  3,  2,
                    -1,  3,  2,  1, -1,
                     1,  2,  1, -1, -3,
                     1, -1,  3, -1,  2), nrow=5, byrow=TRUE)
eg <- eigen(B)
eg$values
max(Mod(eg$values))
```
There are 5 different eigenvalues therefore the eigenspace is full rank (rank is 5). Therefore, $||Bx||_2\leq|\lambda_{max}|$. Note that in this case $|\lambda_{max}|\approx0.299<1$. Therefore, $||B||_2<1$ and subsequently $\sum_{i=0}^kB^i\to(I-B)^{-1}$ and $B^k\to0$ which implies that the algorithm converges.






























